---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eddi: R package for the NOAA Evaporative Demand Drought Index

[![CircleCI](https://circleci.com/gh/earthlab/eddi/tree/master.svg?style=svg)](https://circleci.com/gh/earthlab/eddi/tree/master)
[![Codecov](https://img.shields.io/codecov/c/github/earthlab/eddi.svg)](https://codecov.io/gh/earthlab/eddi)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

The eddi package facilitates access to the NOAA [Evaporative Demand Drought 
Index](https://www.esrl.noaa.gov/psd/eddi/) (EDDI) data product. 

## Installation

You can install the development version of eddi with devtools:

``` r
# install.packages("devtools")
devtools::install_github("earthlab/eddi")
```

## Example

The EDDI product exists for multiple timescales,  including the 1 to 12 week 
and 1 to 12 months scales. 
Shorter time scales can detect short term droughts, e.g., "flash droughts", 
and longer time scales are appropriate for detecting long term drought.
For more information see https://www.esrl.noaa.gov/psd/eddi/.

This is a basic example which shows you how to get EDDI data for Nov 29, 2018 at
the one month timescale:

```{r example}
library(eddi)

eddi_data <- get_eddi(date = "2018-11-29", timescale = "1 month")
eddi_data
```

This will always return a `RasterStack` object with each layer in the stack 
corresponding to a date, that can be visualized using `raster::plot`. 
Here, large positive values indicate exceptionally dry conditions, and 
large negative values indicate exceptionally wet conditions, with values of 0
indicating median EDDI values.

```{r plot-example}
color_pal <- colorRampPalette(c("blue", "lightblue", "white", "pink", "red"))
raster::plot(eddi_data, col = color_pal(255), main = "EDDI data for 2018-11-29")
```

## EDDI Resources 

A user guide for EDDI can be found here: 
https://www.esrl.noaa.gov/psd/eddi/pdf/EDDI_UserGuide_v1.0.pdf

For the science behind EDDI, see these two papers: 

* M. Hobbins, A. Wood, D. McEvoy, J. Huntington, C. Morton, M. Anderson, and 
C. Hain (June 2016): The Evaporative Demand Drought Index: Part I – Linking 
Drought Evolution to Variations in Evaporative Demand. J. Hydrometeor., 
17(6),1745-1761, doi:10.1175/JHM-D-15-0121.1.  
* D. J. McEvoy, J. L. Huntington, M. T. Hobbins, A. Wood, C. Morton, 
M. Anderson, and C. Hain (June 2016): The Evaporative Demand Drought Index: 
Part II – CONUS-wide Assessment Against Common Drought Indicators. 
J. Hydrometeor., 17(6), 1763-1779, doi:10.1175/JHM-D-15-0122.1.
