---
title: "Get EDDI data over a region of interest and save a GeoTIFF"
author: "Maxwell B. Joseph"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create a GeoTIFF for a region of interest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 8, 
  fig.height = 5
)

NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(purl = NOT_CRAN, eval = NOT_CRAN)
```

The eddi R package provides easy access to the Evaporative Demand Drought Index
(EDDI) data - an experimental drought monitoring and early warning guidance tool 
produced by the National Oceanic and Atmospheric Administration. 
EDDI is available at multiple timescales, from weekly to monthly, providing 
insight into short-term flash droughts, and long-term droughts. 
More information on the EDDI product is available on the 
[NOAA EDDI homepage](https://www.esrl.noaa.gov/psd/eddi/).

This vignette covers a common use case for EDDI: acquiring data over a region
of interest defined by a shapefile, masking the EDDI data to that region, 
and saving GeoTIFF files containing EDDI data for the region of interest. 

## Defining a region of interest

By default, the eddi package returns data for the continental United States, 
southern parks of Canada, and northern parts of Mexico. 
But, you may only be interested in a region of interest, as defined by a 
shapefile. 
Here, you will load a shapefile for the state of North Carolina that is 
distributed by default with the [sf](https://cran.r-project.org/package=sf)
package. 

```{r load-roi-shapefile}
library(sf)
library(raster)
library(eddi)

roi <- st_read(system.file("shape/nc.shp", package="sf"))
```

If you are using a different shapefile, replace
`system.file("shape/nc.shp", package="sf")` with its file path, e.g., 
`st_read("path/to/file.shp")`.

The `roi` object contains multiple columns of data, and a `geometry` column
that contains spatial information on the region of interest, which in this 
case consists of multiple counties. 

```{r print-roi-shapefile}
roi
```

Because you don't necessarily care about each county, but rather you want the
entire state (including all counties) you can use a spatial union to join data
from all counties:

```{r}
roi <- st_union(roi)
roi
```


## Acquiring EDDI data

To acquire EDDI data, you can use the `get_eddi()` function. 
You will fetch the 1 week timescale data for July, 2018:

```{r get-eddi-data}
eddi_raster <- get_eddi(date = "2018-07-01", timescale = "1 week")
```

The `eddi_raster` object is a `RasterStack` with one layer, and you can see
information on the spatial extent, resolution, and coordinate reference 
system by printing the object: 

```{r inspect-eddi}
eddi_raster
```

Plot the data with a custom color palette to see what the data look like:

```{r plot-eddi-conus}
color_pal <- colorRampPalette(c("blue", "lightblue", "white", "pink", "red"))
plot(eddi_raster, col = color_pal(255))
```


## Masking to the region of interest

Now you want to subset or mask the EDDI data to the region of interest. 
First, you need to ensure that the raster data and the polygon for the region 
of interest have the same coordinate reference system. 

```{r reproject-to-same-crs}
roi_reprojected <- st_transform(roi, crs = projection(eddi_raster))
```

Now, graphically verify that they align as expected: 

```{r plot-eddi-with-shp}
plot(eddi_raster, col = color_pal(255))
plot(roi_reprojected, add = TRUE)
```

Now, you can crop the EDDI data to match extents with the region of interest, 
then mask the raster set all values outside of the region of interest to `NA`. 
Because the raster package requires sp objects, rather than sf objects, you
will coerce our roi to a sp object first. 

```{r mask-eddi}
roi_sp <- as(roi_reprojected, 'Spatial')
cropped_eddi <- crop(eddi_raster, roi_sp)
masked_eddi <- mask(cropped_eddi, roi_sp)
```

You can plot the masked EDDI raster along with the ROI to confirm:

```{r}
plot(masked_eddi, col = color_pal(255))
plot(roi_sp, add = TRUE)
```

## Saving GeoTIFF output

To write a GeoTIFF file of our `masked_eddi` object, you can use `writeRaster`:

```{r write-tif, eval = FALSE}
writeRaster(masked_eddi, 'eddi-over-roi.tif')
```
